### 2024-09-09 data-visualization.qmd
20240909 - horizontal y-axis title

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    color = "#000000",
    fill = "#0099F8"
  ) +
  labs(
    x = "Fantasy Points",
    title = "Histogram of Fantasy Points"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    fill = position_group)
) +
  geom_density(alpha = 0.7) + # add transparency
  labs(
    x = "Fantasy Points",
    fill = "Position",
    title = "Density Plot of Fantasy Points by Position"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  labs(
    x = "Fantasy Points",
    title = "Histogram of Fantasy Points with Overlaid Density and Rug Plot"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasy_points,
    fill = position_group)
) +
  geom_boxplot() +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Box-and-Whisker Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasy_points,
    fill = position_group)
) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Violin Plot of Fantasy Points by Position",
    subtitle = "Lines represent the 25th, 50th, and 75th quantiles"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    y = position_group,
    group = position_group,
    fill = position_group)
) +
  geom_density_ridges(
    rel_min_height = 0.0085, # remove trailing tails
  ) +
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    y = position_group,
    group = position_group,
    fill = factor(stat(quantile)))
) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = 4,
    quantile_lines = TRUE
  ) +
  # scale_fill_viridis_d() # if want to use viridis color scheme
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot by Position",
    subtitle = "Lines represent the 25th, 50th, and 75th quantiles",
    fill = "Quartiles"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(player_display_name == "Tom Brady"),
  mapping = aes(
    x = season,
    y = fantasy_points
    )
) +
  geom_line(
    linewidth = 1.5,
    color = "blue"
  ) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Bar Plot of Tom Brady's Fantasy Points by Season"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense,
  mapping = aes(
    x = season,
    y = fantasy_points,
    group = player_id,
    color = player_display_name)
) +
  geom_line() +
  gghighlight::gghighlight(
    player_display_name == "Tom Brady",
    label_key = player_display_name) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Fantasy Points by Season and Player",
    subtitle = "(Tom Brady in Red)"
  ) +
  theme_classic()
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary,
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none")
```

@@

ggplot2::ggplot(
  data = epa_combined,
  aes(
    x = off_epa,
    y = def_epa)) +
  nflplotR::geom_mean_lines(
    aes(
      x0 = off_epa ,
      y0 = def_epa)) +
  nflplotR::geom_nfl_logos(
    aes(
      team_abbr = team),
      width = 0.065,
      alpha = 0.7) +
  labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    title = "2023 NFL Offensive and Defensive EPA per Play"
  ) +
  theme_classic() +
  scale_y_reverse()
```

>>>>

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    color = "#000000",
    fill = "#0099F8"
  ) +
  labs(
    x = "Fantasy Points",
    title = "Histogram of Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    fill = position_group)
) +
  geom_density(alpha = 0.7) + # add transparency
  labs(
    x = "Fantasy Points",
    fill = "Position",
    title = "Density Plot of Fantasy Points by Position"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  labs(
    x = "Fantasy Points",
    title = "Histogram of Fantasy Points with Overlaid Density and Rug Plot"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasy_points,
    fill = position_group)
) +
  geom_boxplot() +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Box-and-Whisker Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasy_points,
    fill = position_group)
) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Violin Plot of Fantasy Points by Position",
    subtitle = "Lines represent the 25th, 50th, and 75th quantiles"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    y = position_group,
    group = position_group,
    fill = position_group)
) +
  ggridges::geom_density_ridges(
    rel_min_height = 0.0085, # remove trailing tails
  ) +
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasy_points,
    y = position_group,
    group = position_group,
    fill = factor(after_stat(quantile)))
) +
  ggridges::stat_density_ridges(
    rel_min_height = 0.0085, # remove trailing tails
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = 4,
    quantile_lines = TRUE
  ) +
   scale_fill_viridis_d() + # use viridis color scheme
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot of Fantasy Points by Position",
    subtitle = "Vertical lines represent the 25th, 50th, and 75th quantiles",
    fill = "Quartile"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>%
    filter(player_display_name == "Tom Brady"),
  mapping = aes(
    x = season,
    y = fantasy_points
    )
) +
  geom_line(
    linewidth = 1.5,
    color = "blue"
  ) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Bar Plot of Tom Brady's Fantasy Points by Season"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense,
  mapping = aes(
    x = season,
    y = fantasy_points,
    group = player_id,
    color = player_display_name)
) +
  geom_line() +
  gghighlight::gghighlight(
    player_display_name == "Tom Brady",
    label_key = player_display_name) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Fantasy Points by Season and Player",
    subtitle = "(Tom Brady in Red)"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary,
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group,
    color = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  scale_fill_brewer(palette = "Dark2") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

ggplot2::ggplot(
  data = epa_combined,
  aes(
    x = off_epa,
    y = def_epa)) +
  nflplotR::geom_mean_lines(
    aes(
      x0 = off_epa ,
      y0 = def_epa)) +
  nflplotR::geom_nfl_logos(
    aes(
      team_abbr = team),
      width = 0.065,
      alpha = 0.7) +
  labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    title = "2023 NFL Offensive and Defensive EPA per Play"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
  scale_y_reverse()
```
